# Ggplot Basics {#sec-chapter}

```{r}
#| echo: false
#| message: false
source("_common.R")
```

> ### Learning Objectives {.unnumbered}
>
> * Get
> * Understand
> * Know 
> * Know 
> * Know 

## Reading in Data

Use the `here` package to create file paths

```{r, eval=FALSE}
path <- here::here("folder", "file.csv")
```

Import data with these functions:

File type  | Function       | Library
-----------|----------------|----------
`.csv`     | `read_csv()`   | **readr**
`.txt`     | `read.table()` | **utils**
`.xlsx`    | `read_excel()` | **readxl**

#### Importing Comma Separated Values (.csv) {.unnumbered}

Read in `.csv` files with `read_csv()`:

```{r}
library(tidyverse)
library(here)

csvPath <- here('data', 'milk_production.csv')
milk_production <- read_csv(csvPath) #<<

head(milk_production)
```

#### Importing Text Files (.txt) {.unnumbered}

Read in `.txt` files with `read.table()`:

```{r}
txtPath <- here('data', 'nasa_global_temps.txt')
global_temps <- read.table(txtPath, skip = 5, header = FALSE) #<<

head(global_temps)
```

#### Importing Text Files (.txt) {.unnumbered}

Read in `.txt` files with `read.table()`:

```{r}
txtPath <- here('data', 'nasa_global_temps.txt')
global_temps <- read.table(txtPath, skip = 5, header = FALSE)
names(global_temps) <- c('year', 'no_smoothing', 'loess') # Add header #<<

head(global_temps)
```

# Importing Excel Files (.xlsx) {.unnumbered}

Read in `.xlsx` files with `read_excel()`:

```{r}
library(readxl)

xlsxPath <- here('data', 'pv_cell_production.xlsx')
pv_cells <- read_excel(xlsxPath, sheet = 'Cell Prod by Country', skip = 2) #<<
```

```{r}
glimpse(pv_cells)
```


#### Importing Excel Files (.xlsx) {.unnumbered}

Read in `.xlsx` files with `read_excel()`:

```{r}
library(readxl)

xlsxPath <- here('data', 'pv_cell_production.xlsx')
pv_cells <- read_excel(xlsxPath, sheet = 'Cell Prod by Country', skip = 2) %>%
  mutate(Year = as.numeric(Year)) %>% # Convert "non-years" to NA #<<
  filter(!is.na(Year)) # Drop NA rows in Year #<<
```

```{r}
glimpse(pv_cells)
```
