# Summarizing {#sec-chapter}

```{r}
#| echo: false
#| message: false
#| include: false
source("_common.R")

# Read in data sets
milk_production <- read_csv(here::here('data', 'milk_production.csv'))
lotr_words <- read_csv(here::here('data', 'lotr_words.csv'))
marathon <- read_csv(here::here('data', 'marathon.csv'))
tb_cases <- read_csv(here::here('data', 'tb_cases.csv'))
wildlife_impacts <- read_csv(here::here('data', 'wildlife_impacts.csv'))
daysToShip <- data.frame(
    order = seq(12),
    warehouseA = c(3,3,3,4,4,4,5,5,5,5,5,5),
    warehouseB = c(1,1,1,3,3,4,5,5,5,6,7,10)
)
```


## Types of Data

**Categorical**

Subdivide things into _groups_

- What type?
- Which category?

**Numerical**

Measure things with numbers

- How many?
- How much?

#### Categorical (discrete) variables: {.unnumbered}

**Nominal**

- Order doesn't matter
- Differ in "name" (nominal) only

e.g. `country` in TB case data:

```{r, echo=FALSE}
tb_cases
```

**Ordinal**

- Order matters
- Distance between units not equal

e.g.: `Placement` 2017 Boston marathon:

```{r, echo=FALSE}
marathon %>%
  dplyr::select(Placement = Overall, `Official Time`, Name) %>%
  head()
```

#### Numerical data: {.unnumbered}

**Interval**

- Numerical scale with arbitrary starting point
- No "0" point
- Can't say "x" is double "y"

e.g.: `temp` in Beaver data

```{r, echo=FALSE}
beaver1 %>%
    arrange(day) %>%
    head()
```

**Ratio**

- Has a "0" point
- Can be described as percentages
- Can say "x" is double "y"

e.g.: `height` & `speed` in wildlife impacts

```{r, echo=FALSE}
wildlife_impacts %>%
    filter(!is.na(height), ! is.na(speed)) %>%
    dplyr::select(incident_date, height, speed) %>%
    head()
```

### Key Questions

**Categorical**

<span style="color:blue;">Does this work? Does the order matter?</span>

Yes: **Ordinal**

No: **Nominal**

## Numerical

<span style="color:blue;">Is there a "baseline"?</span>

Yes: **Ratio** 

No: **Interval**

### Variable Encoding
**Be careful of how variables are encoded!**

When numbers are categories

- "Dummy coding": e.g., `passedTest` = `1` or `0`)
- "North", "South", "East", "West" = `1`, `2`, `3`, `4`

When ratio data are discrete (i.e. counts)

- Number of eggs in a carton, heart beats per minute, etc.
- Continuous variables measured discretely (e.g. age)

### Time

- As _ordinal_ categories: "Jan.", "Feb.", "Mar.", etc.
- As _interval_ scale: "Jan. 1", "Jan. 2", "Jan. 3", etc.
- As _ratio_ scale: "30 sec", "60 sec", "70 sec", etc.
